Curve counting in genus one: elliptic singularities and relative geometry

### 2019-06-26 Reduced Gromov-Witten Theory of Hypersurfaces in Genus One.tex
wrote some things

\subsection{Recursive description of the divisors: type $C^0$}

\subsubsection{Restricting the possible combinatorial types} Recall that a boundary divisor $\Dcal \subseteq \Dcal_{1,\alpha,k}(\PP^N|H,d)$ is said to be of \textbf{type $C^0$} if the core is assigned zero degree by the combinatorial type. In this case, the combinatorial type is not as tightly constrained as previously (it need not be a bipartite graph with a single interior vertex). Nevertheless, we have the following lemma:
\begin{lemma} 
Let $\Dcal \subseteq \Dcal_{1,\alpha,k}(\PP^N|H,d)$ be a locus of type $C^0$ and let $\Delta$ be the corresponding combinatorial type. Then for every nonzero vertex $\diamond$ of the expanded target, the fibre of the tropical map over $\diamond$ contains exactly one stable vertex (together with a number of trivial bubbles), and at least one vertex which lies on the radius (which may or may not be stable).
\end{lemma}
\begin{proof}
By stability we always need at least one stable vertex lying over each  vertex of the target. If there was more than one, then (since both the insertions and the factorisation condition are pulled back from the moduli space of maps without expansions) it can be seen, as in the proof of Lemma \ref{lem:combs}, that the contribution vanishes.

@@

%%%%%%%%
\begin{comment}
\begin{center}
\begin{figure}
 \begin{tikzpicture}[ipe import]
  %\useasboundingbox (0, 0) rectangle (595, 842);
  \draw
    (64, 672)
     -- (128, 688)
     -- (64, 704)
     -- (64, 704);
  \draw
    (128, 688)
     -- (64, 736)
     -- (64, 736);
  \draw[red]
    (128, 688) circle[radius=65.9697];
  \draw
    (128, 688)
     -- (112, 752)
     -- (64, 768)
     -- (64, 768);
  \draw
    (112, 752)
     -- (64, 784)
     -- (64, 784);
  \pic
     at (64, 672) {ipe disk};
  \pic
     at (64, 704) {ipe disk};
  \pic
     at (64, 736) {ipe disk};
  \pic
     at (64, 768) {ipe disk};
  \pic
     at (64, 784) {ipe disk};
  \pic
     at (112, 752) {ipe disk};
  \pic
     at (75.2242, 727.5818) {ipe cross};
  \pic
     at (112, 700) {ipe cross};
  \pic
     at (112, 692) {ipe cross};
  \pic
     at (112, 684) {ipe cross};
  \pic
     at (75.8121, 701.047) {ipe cross};
  \pic
     at (75.0705, 674.4385) {ipe cross};
  \filldraw[fill=white,->]
    (64, 608)
     -- (208, 608);
  \pic
     at (64, 608) {ipe disk};
  \draw
    (128, 688)
     -- (208, 688);
  \pic
     at (73.6978, 608.0203) {ipe cross};
  \pic
     at (111.8758, 608.0203) {ipe cross};
  \pic
     at (127.6736, 608.0203) {ipe cross};
  \pic[fill=white]
     at (128, 688) {ipe fdisk};
  \node[ipe node]
     at (86.863, 690.301) {$m$};
  \node[ipe node]
     at (87.521, 671.212) {$m$};
  \node[ipe node]
     at (48.026, 668.579) {$R_1$};
  \node[ipe node]
     at (48.026, 687.668) {$\ldots$};
  \node[ipe node]
     at (47.368, 700.832) {$R_k$};
  \node[ipe node]
     at (48.026, 725.846) {$R_{k+1}$};
  \node[ipe node]
     at (48.026, 784.429) {$R_s$};
  \node[ipe node]
     at (48.026, 766.657) {$\ldots$};
  \node[ipe node]
     at (108.585, 757.441) {$R_r$};
  \node[ipe node]
     at (80, 744.935) {$\ldots$};
  \node[ipe node]
     at (130.965, 677.794) {$E$};
\end{tikzpicture}
\caption{A ray of type $C^0$.}\label{fig:ray_example}
\end{figure}
\end{center}
\end{comment}
%%%%%%%%%%

We observe that the cone $\sigma \in \Sigma^{\Kim}$ corresponding to the stratum $\Ecal \subseteq \ol\Mcal_{1,\alpha}(\PP^N|H,d)$ is isomorphic to $\N^r$ where $r$ is the number of target expansions.
The discrepancy between the dimension of $\sigma$ and the dimension of the ray $\tau$ is to be expected, since the loci of type $C^0$ are contained in the exceptional locus of the blow-up.

We write $\Delta_0$ for the collection of stable vertices of the combinatorial type which are mapped to $0$ by the tropical map, and $\Delta_{>0}$ for the stable vertices which are mapped to $\RR_{>0}$. For each vertex $v$ there is associated discrete data $\Gamma_v$, which defines a moduli space of relative stable maps if $v \in \Delta_0$ and a moduli space of rubber maps if $v \in \Delta_{>0}$. Then $\Ecal$ admits a finite splitting morphism onto the fibre product
\begin{equation*} \Ecal \xrightarrow{\rho} \left( \prod_{v \in \Delta_0} \ol\Mcal_{\Gamma_v}(\PP^N|H) \times \prod_{v \in \Delta_{>0}} \ol\Mcal^{\leftrightarrow}_{\Gamma_v}(P|H_0+H_\infty) \right) \times_{H^{2E}} H^E \end{equation*}
where $E$ is the number of edges of the (stabilised) dual graph, and the fibre product is taken over the appropriate evaluation maps. Note that only the factor corresponding to the core vertex has non-zero genus, and that it must have zero degree.

\begin{remark}As in \S \ref{subsubsection splitting degree} above, we should calculate the degree of $\rho$ and the vanishing order $\lambda_{\Dcal}$ of the section $s_k$ along $\Dcal$. The degree of $\rho$ is given exactly as in Lemma \ref{lem:saturation}, where now we obtain one such factor for each bounded edge of the tropical target. The vanishing order $\lambda_{\Dcal}$ is slightly more delicate, but may be computed as the index of an explicit morphism of one-dimensional lattices, as explained in the proof of Lemma \ref{lemma vanishing order} above.
\end{remark}

Typically, $\Ecal$ will have true dimension larger than that of $\Dcal$. We will study the map $\Dcal \to \Ecal$ in two steps: first, we will describe the map $\widetilde\Dcal \to \Ecal$, and then we will explain how to express the class of the substack $\Dcal \subseteq \widetilde\Dcal$ in terms of tautological classes.

\subsubsection{Constructing $\widetilde\Dcal$ from $\Ecal$}
The geometry of $\widetilde\Dcal\to\Ecal$ is encoded by the combinatorics of the alignment. The reader should bear the following toric construction in mind: suppose we subdivide a fan, and we want to describe the relation between the varieties corresponding respectively to a cone $\rho$ of the subdivision, and the original cone $\sigma$ containing it. The fan of $X_\tau$ is obtained by projecting the star of $\tau$ to the quotient $N_\tau=N_{\RR}/\langle\tau\rangle_{\RR}$ of the ambient vector space by the linear span of the cone\footnote{We find it useful to encode this within the formalism of extended tropicalisations \cite{?}.}. Notice that the dimension of the vector spaces $N_\rho$ and $N_\sigma$ (or the corresponding toric varieties) is off by the difference $c$ between the dimension of the cones. Furthermore, we may need to subdivide the fan of $X_\sigma$ in order for cones of $\Sigma_\rho$ to be mapped surjectively to cones of $\Sigma_\sigma$. The geometric interpretation of this is that $X_\tau$ is obtained as an appropriate compactification of a toric bundle of dimension $c$ over $X_\sigma$; observe that the fiber dimension may not be constant (but it can be made so by blowing up $X_{\sigma}$, see \cite[Lemma 4.1]{AK}). It is a fundamental insight due to Tevelev \cite{Tevelev} that a good compactification of the toric bundle is suggested by the tropicalisation of the locus we are interested in: here, ``good'' means that the intersection of the interesting locus with the boundary is dimensionally transverse. In the case at hand, the interesting locus is that of maps satisfying the factorisation property (i.e. smoothable maps), and the tropicalisation has been worked out in \cite{RSPW2} to consist of well-spaced maps.

We proceed to describe $\Dcal$ step-by-step, building on $\Ecal$.

\subsubsection{The generic alignment} In the first line of the product above, we have grouped the $k$ factors adjacent to the core with maximal expansion factor $m$. It is fundamental to realise that, generically on $\Ecal$, the vertices corresponding to these factors are (equidistant and) already aligned with respect to the core. We can construct a modification $\widetilde\Ecal$ of $\mathcal E$ by:
\begin{enumerate}
 \item radially aligning the core, and imposing the double factorisation condition, thus effectively replacing the factor
 \[\ol\Mcal_{1,\alpha^{(0)}\cup(-m,\ldots,-m,-\mu_{k+1},\ldots,-\mu_r)\cup(-m_i)_{i\in\{k+1,\ldots,s\}\setminus(\bigcup I_j)}}^{\leftrightarrow}(P|H_0+H_\infty,0)\]
 by $\VZ_{1,\alpha^{(0)}\cup(-m,\ldots,-m,-\mu_{k+1},\ldots,-\mu_r)\cup(-m_i)_{i\in\{k+1,\ldots,s\}\setminus(\bigcup I_j)}}^{\leftrightarrow}(P|H_0+H_\infty,0)$ in the product above; in particular, there is a universal cotangent line bundle at the markings $\mathbb L_E$, obtained by suitably twisting any marking cotangent line (they are all isomorphic over a smooth elliptic curve) by the appropriate bubbling loci (see \cite[\S 2.3]{VZ});
 \item aligning the first $k$ factors with respect to the core (this blows up some diagonal subloci of the product, so it is not enough to align the individual factors).
\end{enumerate}
See Figure \ref{fig:fwise_align} below. It is obvious that $\widetilde \Dcal\to \Ecal$ factors through $\widetilde\Ecal$. On $\widetilde\Ecal$, we have a tropical line bundle $\OO(\delta)$ representing the minimum distance from the core to the tropical preimage of $0$; generically, this line bundle coincides with the normal bundle of the locus where the node between the core and one (any) of the first $k$ vertices persists, but on the boundary this needs to be twisted by further smoothing parameters according to the (partial) alignment on $\widetilde\Ecal$. We can obtain a line bundle $\mathbb T_R$ accounting only for the behaviour of the rational side of the fiber product by twisting $\OO(\delta)$ with $\mathbb L_E$.

%%%%%%%%%%%%%
\begin{comment}
\begin{figure}
 \begin{tikzpicture}[ipe stylesheet,scale=.7]
  %\useasboundingbox (0, 0) rectangle (595, 842);
  \pic
     at (129.8488, 703.2932) {ipe disk};
  \pic
     at (97.8488, 719.2932) {ipe disk};
  \pic
     at (97.8488, 687.2932) {ipe disk};
  \draw
    (97.8488, 719.2932)
     -- (129.8488, 703.2932)
     -- (97.8488, 687.2932);
  \draw
    (129.8488, 703.2932)
     -- (161.8488, 687.2932);

  \draw
    (193.8488, 671.2932)
     -- (225.8488, 655.2932);
  \draw
    (225.8488, 655.2932)
     -- (193.8488, 639.2932);
    \draw[ipe dash dashed]
   (193.8488, 671)
    -- (161.8488, 687);
  \draw[ipe dash dashed]
    (193.8488, 639.2932)
     -- (161.8488, 623.2932);
  \draw
    (161.8488, 623.2932)
     -- (129.8488, 607.2932);
  \pic
     at (146.637, 615.0615) {ipe disk};
  \draw
    (129.8488, 607.2932)
     -- (97.8488, 591.2932)
     -- (97.8488, 591.2932);
  \draw
    (147.2948, 615.0622)
     -- (97.2686, 638.1002)
     -- (97.2686, 638.1002);
  \pic
     at (96.6103, 637.4422) {ipe disk};
  \pic
     at (97.9268, 590.7072) {ipe disk};
  \pic
     at (130.1808, 622.9607) {ipe cross};
  \pic
     at (130.1808, 607.1635) {ipe cross};
  \pic
     at (145.9788, 695.3672) {ipe cross};
  \draw[-ipe linear]
    (97.8488, 575.2932)
     -- (257.8488, 575.2932);
  \pic
     at (97.8488, 575.2932) {ipe disk};
  \pic
     at (129.8488, 575.2932) {ipe cross};
  \pic
     at (145.8488, 575.2932) {ipe cross};
  \pic
     at (225.8488, 575.2932) {ipe cross};
  \draw
    (129.8488, 703.2932)
     -- (257.8488, 703.2932);
  \draw
    (225.8488, 655.2932)
     -- (257.8488, 655.2932)
     -- (257.8488, 655.2932);
  \draw
    (146.3018, 614.861)
     -- (258.5378, 614.404);
  \node[ipe node]
     at (189.423, 676.278) {$m$};
  \node[ipe node]
     at (189.094, 629.543) {$m$};
  \node[ipe node]
     at (111.421, 715.773) {$m_1$};
  \node[ipe node]
     at (110.104, 590.707) {$m_2$};
  \node[ipe node]
     at (110.763, 564.378) {$\ell_1$};
  \node[ipe node]
     at (133.801, 564.377) {$\ell_2$};
  \node[ipe node]
     at (146.308, 543.972) {$\frac{\ell_1}{m_1}+\frac{\ell_2}{m} \lesseqgtr \frac{\ell_1+\ell_2}{m_2}$};
  \pic
     at (473.6651, 658.9228) {ipe disk};
  \pic
     at (353.8488, 719.2932) {ipe disk};
  \pic
     at (353.8488, 687.2932) {ipe disk};
  \draw
    (353.8488, 719.2932)
     -- (385.8488, 703.2932)
     -- (353.8488, 687.2932);
  \draw
    (386.7731, 703.2932)
     -- (418.7731, 687.2932);
  \draw[ipe dash dashed]
    (417.8488, 687.2932)
     -- (449.8488, 671.2932)
     -- (449.8488, 671.2932);

  \draw
    (449.8488, 671.2932)
     -- (481.8488, 655.2932)
     -- (481.8488, 655.2932);
  \draw
    (481.8488, 655.2932)
     -- (449.8488, 639.2932)
     -- (449.8488, 639.2932);
  \draw[ipe dash dashed]
    (449.8488, 639.2932)
     -- (417.8488, 623.2932);
  \draw
    (417.8488, 623.2932)
     -- (385.8488, 607.2932);
  \pic
     at (457.1756, 642.7929) {ipe disk};
  \draw
    (385.8488, 607.2932)
     -- (353.8488, 591.2932)
     -- (353.8488, 591.2932);
  \pic
     at (352.6103, 622.6521) {ipe disk};
  \pic
     at (353.9268, 590.7072) {ipe disk};
  \pic
     at (458.2826, 667.3311) {ipe cross};
  \pic
     at (459.207, 575.7344) {ipe cross};
  \draw[-ipe linear]
    (353.8488, 575.2932)
     -- (513.8488, 575.2932);
  \pic
     at (353.8488, 575.2932) {ipe disk};
  \pic
     at (385.8488, 575.2932) {ipe cross};
  \pic
     at (475.7994, 575.2932) {ipe cross};
  \pic
     at (481.8488, 575.2932) {ipe cross};
  \draw[shift={(472.741, 659.847)}, xscale=0.3271, yscale=-1.7893]
    (0, 0)
     -- (128, 0);
  \draw
    (481.8488, 655.2932)
     -- (513.8488, 655.2932)
     -- (513.8488, 655.2932);
  \draw[shift={(459.26, 642.878)}, xscale=0.4834, yscale=0.4807]
    (0, 0)
     -- (112.236, -0.457);
  \draw[red]
    (482.3765, 654.0394) circle[radius=9.9987];
  \draw
    (354.8117, 621.686)
     -- (386.2407, 606.8958);
  \pic
     at (386.2407, 606.8958) {ipe disk};
  \pic
     at (386.2407, 703.956) {ipe disk};
  \draw[red]
    (360.3506, 584.6012)
     arc[start angle=160.4967, end angle=217.0656, x radius=120.4239, y radius=-143.3401];
  \pic
     at (473.9974, 650.6919) {ipe cross};
  \pic
     at (364.92, 693.131) {ipe cross};
  \pic
     at (372.3151, 710.6943) {ipe cross};
  \pic
     at (359.3737, 593.2977) {ipe cross};
  \draw
    (226.3225, 655.8881)
     -- (258.6759, 665.1319);
  \draw
    (226.3225, 654.0394)
     -- (258.6759, 646.6443);
  \draw[-ipe linear]
    (271.6173, 766.814)
     -- (337.2485, 767.7384);

  \pic
     at (272.5417, 782.5285) {ipe disk};
  \pic
     at (271.6173, 815.8063) {ipe disk};
  \draw
    (305.8194, 798.2431)
     -- (272.5417, 815.8063);
  \draw
    (305.8194, 798.2431)
     -- (272.5417, 782.5285)
     -- (272.5417, 782.5285);
  \draw
    (272.5417, 782.5285)
     -- (336.3241, 782.5285);
  \draw
    (272.5417, 815.8063)
     -- (336.3241, 814.8819);
  \draw
    (304.8951, 799.1674)
     -- (335.3997, 799.1674);
  \draw
    (305.8194, 799.1674)
     -- (336.3241, 804.7137);
  \draw
    (305.8194, 798.2431)
     -- (335.3997, 792.6968);
  \draw
    (387.1651, 703.956)
     -- (511.0324, 703.956)
     -- (511.0324, 703.0316);
  \draw
    (388.0895, 605.0471)
     -- (511.9568, 605.0471);
  \pic
     at (272.5417, 766.814) {ipe disk};
  \draw[red]
    (304.8951, 797.3187) circle[radius=35.9679];

    \pic[fill=white]
     at (225.8488, 655.2932) {ipe fdisk};

     \pic[fill=white]
     at (481.8488, 655.2932) {ipe fdisk};

     \pic[fill=white]
     at (304.8951, 799.1674) {ipe fdisk};
\end{tikzpicture}
\caption{Two degenerations of the combinatorial type above, which represents a ray. On the left-hand side, the alignment holds only generically. On the right-hand side, it is not enough to align each individual factor.}\label{fig:fwise_align}
\end{figure}
\end{comment}
%%%%%%%%%%%%%%%%%%

\subsubsection{The projective bundle}\label{subs:proj_bun} Generically on $\widetilde\Ecal$, the compactified torus bundle $\widetilde \Dcal\to \widetilde\Ecal$ coincides with the following projective bundle:
\[\mathcal P=\PP_{\widetilde\Ecal}(\mathbb T_R\oplus\bigoplus_{j=k+1}^r \mathbb T_{q_j}R_j),\]
where $\mathbb T_{q_j}R_J$ is the tangent line at the gluing marking closest to the core on the $j$-th factor of the product in the second line of \eqref{eq:product}. The open stratum in the fiber consists of moduli of attachments (i.e. moduli for the contraction to the Smyth's singularity); the compactification is by adding strictly semistable Gorenstein singularities as in \cite[\S\S 2.2-2.3]{SMY2}. In terms of alignment, a coordinate hyperplane corresponds to pushing the corresponding vertex off the circle, and destabilising the corresponding edge to the core. Tropically, the fibers can be analysed as follows: pushing some of the vertices off the circle creates new finite length edges with coordinates $\ell_j\in\RR_{\geq_0}$ (see Figure \ref{fig:off_we_go}); not all of them can be strictly positive, because the circle needs to pass through at least one vertex of positive horizontal degree. This is patently the fan of a projective space.

%%%%%%%%%%%%%%%%
\begin{comment}
\begin{figure}
\begin{tikzpicture}[ipe import,scale=.8]
  %\useasboundingbox (0, 0) rectangle (595, 842);
  \draw
    (64, 672)
     -- (128, 688)
     -- (64, 704)
     -- (64, 704);
  \draw
    (128, 688)
     -- (64, 736)
     -- (64, 736);
  \draw[red]
    (128, 688) circle[radius=65.9697];
  \draw
    (128, 688)
     -- (112, 752)
     -- (64, 768)
     -- (64, 768);
  \draw
    (112, 752)
     -- (64, 784)
     -- (64, 784);
  \pic
     at (64, 672) {ipe disk};
  \pic
     at (64, 704) {ipe disk};
  \pic
     at (64, 736) {ipe disk};
  \pic
     at (64, 768) {ipe disk};
  \pic
     at (64, 784) {ipe disk};
  \pic
     at (112, 752) {ipe cross};
  \pic
     at (75.2242, 727.582) {ipe cross};
  \pic
     at (112, 700) {ipe cross};
  \pic
     at (112, 692) {ipe cross};
  \pic
     at (112, 684) {ipe cross};
  \pic
     at (75.8121, 701.047) {ipe cross};
  \pic
     at (75.0704, 674.4388) {ipe cross};
  \filldraw[fill=white]
    (64, 608)
     -- (208, 608)
     -- (208, 608);
  \pic
     at (64, 608) {ipe disk};
  \draw
    (128, 688)
     -- (208, 688)
     -- (208, 688);
  \pic
     at (73.6978, 608.02) {ipe cross};
  \pic
     at (111.876, 608.0202) {ipe cross};
  \pic
     at (127.674, 608.0205) {ipe cross};
  \pic[fill=white]
     at (128, 688) {ipe fdisk};
  \node[ipe node]
     at (86.863, 690.301) {$m$};
  \node[ipe node]
     at (87.521, 671.212) {$m$};
  \draw
    (128, 688)
     -- (91.4703, 741.643);
  \draw
    (128, 688)
     -- (160.586, 745.593)
     -- (160.586, 745.593);
  \pic
     at (91.4703, 741.643) {ipe disk};
  \pic
     at (160.486, 745.417) {ipe disk};
  \draw[shift={(91.47, 741.051)}, xscale=0.95, yscale=0.6996]
    (0, 0)
     -- (-26.9879, 7.241)
     -- (-26.9879, 7.241);
  \draw
    (160.486, 745.417)
     -- (64.4824, 799.569);
  \draw
    (160.486, 745.417)
     -- (208.637, 745.593);
  \draw
    (92.1286, 740.985)
     -- (64.4824, 754.15);
  \pic
     at (64.4824, 746.909) {ipe disk};
  \pic
     at (64.4824, 798.91) {ipe disk};
  \pic
     at (64.4824, 753.492) {ipe disk};
  \draw
    (289.118, 672.0005)
     -- (353.118, 688.0005)
     -- (289.118, 704.0005)
     -- (289.118, 704.0005);
  \draw
    (353.118, 688.0005)
     -- (289.118, 736.0005)
     -- (289.118, 736.0005);
  \draw[red]
    (353.118, 688.0005) circle[radius=65.9697];
  \draw
    (353.118, 688.0005)
     -- (337.118, 752.0005)
     -- (289.118, 768.0005)
     -- (289.118, 768.0005);
  \draw
    (337.118, 752.0005)
     -- (289.118, 784.0005)
     -- (289.118, 784.0005);
  \pic
     at (289.118, 672.0005) {ipe disk};
  \pic
     at (289.118, 704.0005) {ipe disk};
  \pic
     at (289.118, 736.0005) {ipe disk};
  \pic
     at (289.118, 768.0005) {ipe disk};
  \pic
     at (289.118, 784.0005) {ipe disk};
  \pic
     at (337.118, 752.0005) {ipe cross};
  \pic
     at (300.3422, 727.5825) {ipe cross};
  \pic
     at (337.118, 700.0005) {ipe cross};
  \pic
     at (337.118, 692.0005) {ipe cross};
  \pic
     at (337.118, 684.0005) {ipe cross};
  \pic
     at (300.9301, 701.0475) {ipe cross};
  \pic
     at (300.1889, 674.4392) {ipe cross};
  \filldraw[fill=white]
    (289.118, 608.0005)
     -- (433.118, 608.0005)
     -- (433.118, 608.0005);
  \pic
     at (289.118, 608.0005) {ipe disk};
  \draw
    (353.118, 688.0005)
     -- (433.118, 688.0005)
     -- (433.118, 688.0005);
  \pic
     at (298.8158, 608.0205) {ipe cross};
  \pic
     at (336.994, 608.0207) {ipe cross};
  \pic
     at (352.792, 608.0209) {ipe cross};
  \pic[fill=white]
     at (353.118, 688.0005) {ipe fdisk};
  \node[ipe node]
     at (311.98, 690.301) {$m$};
  \node[ipe node]
     at (312.639, 671.212) {$m$};
  \draw[shift={(353.118, 687.999)}, xscale=1.1287, yscale=1.1615]
    (0, 0)
     -- (-36.5297, 53.643);
  \draw[shift={(353.119, 687.997)}, xscale=1.4269, yscale=1.3567]
    (0, 0)
     -- (32.586, 57.593)
     -- (32.586, 57.593);
  \pic
     at (310.6643, 751.5171) {ipe disk};
  \pic
     at (400.086, 767.1394) {ipe disk};
  \draw[shift={(311.745, 750.197)}, xscale=0.8205, yscale=-0.5452]
    (0, 0)
     -- (-26.9879, 7.241)
     -- (-26.9879, 7.241);
  \draw[shift={(399.543, 767.54)}, xscale=1.1319, yscale=0.5704]
    (0, 0)
     -- (-96.0036, 54.152);
  \draw[shift={(400.744, 767.146)}, xscale=0.6446, yscale=-1.3688]
    (0, 0)
     -- (48.151, 0.176);
  \draw[shift={(311.717, 751.113)}, xscale=0.8, yscale=0.1453]
    (0, 0)
     -- (-27.6462, 13.165);
  \pic
     at (289.6004, 746.9095) {ipe disk};
  \pic
     at (289.6004, 798.9105) {ipe disk};
  \pic
     at (289.6004, 753.4925) {ipe disk};
  \pic
     at (387.02, 744.935) {ipe cross};
  \pic
     at (317.246, 742.961) {ipe cross};
  \node[ipe node]
     at (315.273, 748.884) {$\ell_{j_1}$};
  \node[ipe node]
     at (397.553, 752.175) {$\ell_{j_2}$};
\end{tikzpicture}
\caption{Moving vertices off the circle creates new tropical parameters.}\label{fig:off_we_go}
\end{figure}
\end{comment}
%%%%%%%%%%%%%

\subsubsection{Factorisation: obtaining $\Dcal$ from $\widetilde\Dcal$} Assume $m\geq 2$. Over $\mathcal P$ we have a tautological sequence:
\[\OO_{F}(-1)\hookrightarrow p^*(F)\xrightarrow{+\operatorname{d}f}\operatorname{ev}_E^*(TH),\]
where $p\colon\mathcal P=\PP(F)\to\widetilde\Ecal$, and $+\operatorname{d}f$ should be interpreted as follows: $k$ copies of $\mathbb T_R$ are taken to account - via the preferred isomorphisms - for all factors of the first product in \eqref{eq:product}; then, the differential of the universal map to $\PP^N$ is applied to each tangent vector (possibly including some of the contracted rational bubbles in the boundary), and the sum of these contributions is taken in $T_{f(E)}\PP^N$ - it actually lies in $T_{f(E)}H$ because all the normal contributions are trivial by assumption. The composition is a section $s$ of $G=\OO_F(1)\otimes \operatorname{ev}_E^*(TH)$, and it vanishes precisely where $f$ factors through the Smyth's singularity determined by the point in the fiber of $p$. Generically on $\widetilde\Ecal$, this is exactly what we are after.

\subsubsection{Dimensional transversality} In general, the vanishing locus of $s$ corresponds to maps $f\colon C\to\PP^N$ and diagrams $C\leftarrow \tilde C\to \bar C$ - with $\bar C$ the elliptic $r$-fold point determined by the position in the fiber of $p\colon\mathcal P\to\widetilde \Ecal$ - such that $f$ factors through $\bar C$. When the curve is allowed to degenerate (so, on the boundary of $\widetilde \Ecal$), this is not always what we want: indeed, due do the combined action of degeneration (in the base direction) and destabilisation (in the fiber direction), it may very well be that $f$ is constant along every component of the elliptic singularity, so that it factors trivially; in other words, the vanishing locus of $s$ is not dimensionally transverse to the boundary of $\mathcal P$.
\begin{lemma}
 For every $K\subseteq\{1,\ldots,r\}$ and $K^c=\{1,\ldots,r\}\setminus K$, $s$ vanishes along the loci where the factors corresponding to $K$ in $\widetilde \Ecal$ degenerate so that $f_k\colon R_k\to\PP^N$ is constant in a neighbourhood of $q_k$, and the fiber coordinates $(\lambda_{h})_{h\in K^c}$ along $p\colon\mathcal P\to\widetilde \Ecal$ are set to be $0$.
\end{lemma}
Call these loci $\Delta_i,i\in I$. See \cite[\S 3.2]{VZ} for a treatment of the analogous situation in the non-relative case; the study of the normal bundles is also dealt with there.

>>>>

\subsection{Recursive description of the divisors: type IIIb} The treatment of the type I, II and IIIa cases above is a fairly straightforward extension of the ideas of Vakil and Gathmann (albeit updated with logarithmic and tropical technology, which streamlines the presentation and simplifies the proofs). The remaining type IIIb case, however, is entirely new. In this section we provide a recursive description of these boundary divisors; this constitutes the technical heart of the paper.

\subsubsection{Description of the possible combinatorial types} Recall that a boundary divisor $\Dcal \subseteq \Dcal_{1,\alpha,k}(\PP^N|H,d)$ is said to be of type IIIb if the core is assigned degree zero by the combinatorial type (this is the only case which we did not treat in Lemma \ref{lem:combs}). In this case, the combinatorial type is not as tightly constrained as before; the target may expand multiple times, and as such the combinatorial type need not be a bipartite graph with a single interior vertex. Nevertheless, we have the following lemma:
\begin{lemma} 
Let $\Dcal$ be a locus of type IIIb and let $\Delta$ be the corresponding combinatorial type. Then for every nonzero vertex $\diamond$ of the expanded tropical target, the fibre of the tropical map over $\diamond$ contains:
\begin{enumerate}
\item exactly one stable vertex (together with a number of trivial bubbles);
\item at least one vertex which lies on the radius (which may or may not be stable).
\end{enumerate}
\end{lemma}
\begin{proof}
By stability, we must always have at least one stable vertex lying over each  vertex of the target. If there was more than one, then (since both the insertions and the factorisation condition are pulled back from the moduli space of maps without expansions) we can show, as in the proof of Lemma \ref{lem:combs}, that the contribution vanishes.

@@

\subsubsection{Recursive description of $\Ecal$}
Similarly to before, we let $\tau \in \Sigma$ be the ray corresponding to $\Dcal \subseteq \VZ_{1,\alpha}(\PP^N|H,d)$, $\sigma \in \Sigma^{\Kim}$ be the minimal cone containing $\tau$ and $\Ecal \subseteq \ol\Mcal_{1,\alpha}(\PP^N|H,d)$ be the stratum corresponding to $\sigma$, so that we have a map $\Dcal \to \Ecal$ (and $\Ecal$ is the minimal stratum through which the map from $\Dcal$ factors). Recall that we also have a unique stratum $\widetilde\Dcal \subseteq \widetilde\VZ_{1,\alpha}(\PP^N|H,d)$ such that $\Dcal = \widetilde\Dcal \cap \VZ_{1,\alpha}(\PP^N|H,d)$. Note that the cone $\sigma$ has dimension equal to the number of levels of the target expansion in the combinatorial type, which in this case may well be strictly larger than $1$.

We can describe the locus $\Ecal$ in a simple recursive manner. We write $\Delta_0$ for the collection of stable vertices of the combinatorial type which are mapped to $0$ by the tropical map, and $\Delta_{>0}$ for the stable vertices which are mapped to $\RR_{>0}$. For each vertex $v$ there is associated discrete data $\Gamma_v$, which defines a moduli space of relative stable maps if $v \in \Delta_0$ and a moduli space of rubber maps if $v \in \Delta_{>0}$. Then $\Ecal$ admits a finite splitting morphism onto the fibre product
\begin{equation}\label{IIIa fibre product} \Ecal \xrightarrow{\rho} \left( \prod_{v \in \Delta_0} \ol\Mcal_{\Gamma_v}(\PP^N|H) \times \prod_{v \in \Delta_{>0}} \ol\Mcal^{\leftrightarrow}_{\Gamma_v}(P|H_0+H_\infty) \right) \times_{H^{2\epsilon}} H^{\epsilon} \end{equation}
where $\epsilon$ is the number of edges of the (stabilised) dual graph, and the fibre product is taken over the appropriate evaluation maps. Thus, tautological integrals over $\Ecal$ are completely determined in terms of integrals over moduli spaces with ``smaller'' combinatorial data.

Notice that the actual dimension of $\Ecal$ will typically be larger than that of $\Dcal$, since $\Ecal$ is contained inside the excess components of the moduli space. We will describe $\Dcal$ in terms of the map $\Dcal \to \Ecal$, proceeding in two steps:
\begin{enumerate}
\item obtaining $\widetilde\Dcal$ from $\Ecal$;
\item expressing the class of $\Dcal \subseteq \widetilde\Dcal$ in terms of tautological classes.
\end{enumerate}

\begin{remark}[Splitting degree and vanishing order, revisited] Similarly to \S \ref{subsubsection splitting degree} above, we can calculate the degree of $\rho$ and the vanishing order $\lambda_{\Dcal}$ of the section $s_k$ along $\Dcal$. The degree of $\rho$ is given exactly as in Lemma \ref{lem:saturation}, where now we obtain one such factor for each bounded edge of the tropical target. The vanishing order $\lambda_{\Dcal}$ is slightly more delicate, but may be computed as the index of an explicit morphism of one-dimensional lattices, as explained in the proof of Lemma \ref{lemma vanishing order}.\end{remark}

\subsubsection{$\widetilde\Dcal$ from $\Ecal$} Recall that the map $\widetilde\Dcal \to \Ecal$ is obtained by restricting the logarithmic blow-up
\begin{equation*} \widetilde\VZ_{1,\alpha}(\PP^N|H,d) \to \ol\Mcal_{1,\alpha}(\PP^N|H,d)\end{equation*}
to the locus $\widetilde\Dcal$. In toric geometry, the method for studying this restriction is as follows. The fan of $\widetilde\Dcal$ is obtained by projecting the star of the ray $\tau$ onto the quotient lattice, and similarly for $\Ecal$. The map $\widetilde\Dcal \to \Ecal$ is then a toric morphism, given by the obvious map of fans. In order to describe this morphism, we first subdivide the fan of $\Ecal$ so that every cone in the fan of $\widetilde\Dcal$ maps surjectively onto a cone. This amounts to taking a toric blow-up of $\Ecal$ which we denote by $\widetilde\Ecal$. There is a natural morphism $\widetilde\Dcal \to \widetilde\Ecal$ which is now flat, and this may be described as a compactification of a torus bundle on $\widetilde\Ecal$.

In the logarithmic setting, we proceed as follows. The stratum $\widetilde\Dcal$ has a natural \emph{reduced} logarithmic structure, obtained by restricting the logarithmic structure on $\widetilde\VZ_{1,\alpha}(\PP^N|H,d)$ and then quotienting out by the normal direction to $\widetilde\Dcal$. Similarly, $\Ecal$ has a natural reduced logarithmic structure, which coincides with the pull-back of the product logarithmic structure on the fibre product \eqref{IIIa fibre product}.

We would like to describe the tropicalisations of these logarithmic stacks. In fact, it is more convenient to describe the \emph{extended tropicalisations} \cite{}. Let $l_1,\ldots,l_m$ denote the edge lengths of the tropical target, forming co-ordinates on the cone $\sigma$. Then the extended tropicalisation of $\Ecal$ is obtained by intersecting the star of $\sigma$ with the face $l_1=\ldots=l_m=\infty$.

For $\widetilde\Dcal$ we have on $\tau$ edge lengths $e_1=\ldots=e_r=e$ separating the core from the vertices on the radius. The extended tropicalisation of $\widetilde\Dcal$ is obtained by intersecting the star of $\tau$ with the face $e=\infty$. Since $l_1=\ldots=l_m=\infty$ implies $e=\infty$ there is certainly a morphism between these cone complexes, and it is clear this morphism is a subdivsion.

Moreover we may describe the subdivision explicitly.

\subsubsection{$\Dcal$ from $\widetilde\Dcal$}
